Models of Lambek calculus with subexponentials

### 2019-06-16 PaperDraft.tex
updates in the draft

@@

@@

  $\Sigma = \langle \mathcal{I}, \preceq, \mathcal{W}, \mathcal{C}, \mathcal{E} \rangle$,
\end{defin}

where $I = \{ s_1, \dots, s_n\}$, $\mathcal{I}, \preceq \rangle$ is a preorder.
$W, C, E$ are upwardly closed subsets of $I$ and $\mathcal{W} \cap \mathcal{C} \subseteq \mathcal{E}$.

@@

    \begin{prooftree}
      \AxiomC{$\Gamma, A, B, \Delta \rightarrow C$}
      \RightLabel{$\bullet \rightarrow$}
      \UnaryInfC{$\Gamma, A \bullet B, \Delta \rightarrow C$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma, A_i, \Delta \rightarrow B$}
      \RightLabel{$\&, i = 1,2 \rightarrow$}
      \UnaryInfC{$\Gamma, A_1 \& A_2, \Delta \rightarrow B$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma, A, \Delta \rightarrow C$}
      \AxiomC{$\Gamma, B, \Delta \rightarrow C$}
      \RightLabel{$\vee \rightarrow$}
      \BinaryInfC{$\Gamma, A \vee B, \Delta \rightarrow C$}
    \end{prooftree}

    \begin{prooftree}
    \AxiomC{$\Gamma, A, \Delta \rightarrow C$}
    \RightLabel{$! \rightarrow$}
    \UnaryInfC{$\Gamma, !^{s} A, \Delta \rightarrow C$}
    \end{prooftree}
  \end{flushleft}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \begin{flushright}
    \begin{prooftree}
      \AxiomC{$A, \Pi \rightarrow B$}
      \RightLabel{$\rightarrow \backslash$}
      \UnaryInfC{$\Pi \rightarrow A \backslash B$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Pi, A \rightarrow B$}
      \RightLabel{$\rightarrow /$}
      \UnaryInfC{$\Pi \rightarrow B / A$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma \rightarrow A$}
      \AxiomC{$\Delta \rightarrow B$}
      \RightLabel{$\rightarrow \bullet$}
      \BinaryInfC{$\Gamma, \Delta \rightarrow A \bullet B$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma \rightarrow A$}
      \AxiomC{$\Gamma \rightarrow B$}
      \RightLabel{$\rightarrow \&$}
      \BinaryInfC{$\Gamma \rightarrow A \& B$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma \rightarrow A_i$}
      \RightLabel{$\rightarrow \vee, i = 1,2$}
      \UnaryInfC{$\Gamma \rightarrow A_1 \vee A_2$}
    \end{prooftree}

    \begin{prooftree}
    \AxiomC{$!^{s_1} A_1, \dots, !^{s_n} A_n \rightarrow A$}
    \RightLabel{$\rightarrow !, \forall j, s_j \succeq s$}
    \UnaryInfC{$!^{s_1} A_1, \dots, !^{s_n} A_n \rightarrow !^{s} A$}
    \end{prooftree}
  \end{flushright}
\end{minipage}
Structural rules:

\begin{minipage}{0.5\textwidth}
  \begin{flushleft}
        \begin{prooftree}
        \AxiomC{$\Gamma, !^{s} A, \Delta, !^{s} A, \Theta \rightarrow B$}
        \RightLabel{${\bf ncontr}_1, s \in C$}
        \UnaryInfC{$\Gamma, !^{s} A, \Delta, \Theta \rightarrow B$}
        \end{prooftree}

@@

@@

  \end{flushleft}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \begin{flushright}

      \begin{prooftree}
        \AxiomC{$\Gamma, !^{s} A, \Delta, \Theta \rightarrow B$}
        \RightLabel{${\bf ex}_2, s \in E$}
        \UnaryInfC{$\Gamma, \Delta, !^{s} A, \Theta \rightarrow A$}
      \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow A$}
    \AxiomC{$\Pi, A, \Delta \rightarrow B$}
    \RightLabel{${\bf cut}$}
    \BinaryInfC{$\Gamma, \Pi, \Delta \rightarrow B$}
  \end{prooftree}
  \end{flushright}
\end{minipage}
\end{defin}

@@

\section{Semantics}

@@

  A quantale is a triple $\mathcal{Q} = \langle A, \bigvee, \cdot \rangle$, such that $\langle A, \bigvee \rangle$
is a complete lattice and $\langle A, \cdot \rangle$ is a semigroup, such that for all indexing set $I$:

@@

Some example of quantales:

@@

@@

  If $\mathcal{Q}_1$, $\mathcal{Q}_2$ are unital quantales, then a unital homomorphism if a quantale homomorphism, such that $f(\varepsilon) = \varepsilon$.
\end{defin}

@@

  Let $\mathcal{Q} = \langle A, \bigvee, \cdot \rangle$ be a quantale. $\mathcal{S} \subseteq \mathcal{Q}$ is said to be a subquantale, if
  $\mathcal{S}$ is closed under multiplication and sups.
\end{defin}

@@

Then, if $f : \mathcal{Q}_1 \to \mathcal{Q}_2$ is a quantale homomorphism, then $f(\mathcal{S}) \subseteq \mathcal{Q}_2$ is a subquantale of $\mathcal{Q}_2$.

In other words, a homomorphic image of subquantale is a subquantale.
\end{prop}

@@

  It is clearly that $f(\mathcal{S}) \subseteq \mathcal{Q}_2$ is a submonoid of $\mathcal{Q}_2$. Let $a_i \in S$ for each $i
  \in I$, so $\bigvee \limits_{i \in I} a_i \in S$, but $f(a_i) \in f(\mathcal{S})$ for any $i \in I$, so $f(\bigvee
  \limits_{i \in I} a_i) = \bigvee \limits_{i \in I} (f (a)) \in f(\mathcal{S})$, so $f(\mathcal{S})$ is closed
  under joins, so $f(\mathcal{S})$ is a subquantale of $\mathcal{Q}_2$
\end{proof}

@@

@@

  An open modality (or quantic conucleus) on quantale $\mathcal{Q}$ is a map $\Box : \mathcal{Q} \to \mathcal{Q}$, such that

\begin{enumerate}
  \item $\Box x \leq x$;
  \item $\Box x = \Box \Box x$;
  \item $x \leq y \Rightarrow \Box x \leq \Box y$;
  \item $\Box x \cdot \Box y = \Box (\Box x \cdot \Box y)$.
\end{enumerate}

For unital quantale, we require that $\Box e = e$.
\end{defin}

Note that, we may replace the last condition on equivalent condition $\Box (x) \cdot \Box (y) \leq \Box (x \cdot y)$.

@@

  We define a partial order on open modalities on $\mathcal{Q}$ as $\Box_1 \leq \Box_2 \Leftrightarrow \forall a \in Q, \Box_1 (a) \leq \Box_2 (a)$.
\end{defin}

\begin{lemma}
  $\Box_1 a_1 \cdot \Box_2 a_2 \leq \Box (\Box_1 a_1 \cdot \Box_2 a_2)$, where $\Box_i \leq \Box, i = 1,2$.
\end{lemma}

@@

  $\begin{array}{lll}
  &\Box_1 a_1 \cdot \Box_2 a_2 \leq & \\
  &\Box_1 (\Box_1 a_1) \cdot \Box_2 (\Box_2 a_2) \leq & \\
  &\Box (\Box_1 a_1) \cdot \Box (\Box_2 a_2) \leq & \\
  &\Box (\Box_1 (a_1) \cdot \Box_2 (a_2))&
  \end{array}$
\end{proof}

\begin{defin}
$ $
\begin{enumerate}
  \item An open modality is called central, if for all
  $a, b \in \mathcal{Q}, \Box a \cdot b = b \cdot \Box a$.

  \item An open modality is called weak square-increasing,
  if for all $a, b \in \mathcal{Q}, \Box a \cdot b \leq \Box a \cdot b \cdot \Box a$ and
  $b \cdot \Box a \leq \Box a \cdot b \cdot \Box a$.

  \item An open modality is called unital, if $\forall a \in Q, \Box a \leq e$.
\end{enumerate}
\end{defin}

@@

  Let $\Box$ be an open modality on some unital quantale $\mathcal{Q} = \langle A, \bigvee, \cdot, e \rangle$.
  Then, if $\Box$ is unital and weak idempotent, then $\Box$ is central.
\end{lemma}

@@

  $\begin{array}{lll}
  & b \cdot \Box a \leq & \\
  & \:\:\:\: \text{Right weak square-increase}& \\
  &\Box a \cdot b \cdot \Box a \leq & \\
  & \:\:\:\: \text{Unitality}& \\
  & \Box a \cdot b \cdot e \leq & \\
  & \:\:\:\: \text{Identity}& \\
  &\Box a \cdot b \leq & \\
  & \:\:\:\: \text{Left weak square-increase}& \\
  &\Box a \cdot b \cdot \Box a \leq & \\
  & \:\:\:\: \text{Unitality}& \\
  &e \cdot b \cdot \Box a \leq & \\
  & \:\:\:\: \text{Identity}& \\
  &b \cdot \Box a&
  \end{array}$

Hence, $b \cdot \Box a = \Box a \cdot b$, so for all $a \in \mathcal Q, \Box a \in \mathcal{Z}(\mathcal{Q})$.

@@

  Let $\mathcal{Q}$ be a quantale and $S \subseteq \mathcal{Q}$ a subquantale,
  then $\Box : \mathcal{Q} \to \mathcal{Q}$, such that $\Box a = \bigvee \{ s \in S \: | \: s \leq a \}$, is an open modality.
\end{prop}

@@

@@

  Let $\Sigma = \langle I, \preceq, W, C, E \rangle$ be a subexponential signature, where $|I| = n$ and
  $\Box_{\mathcal{Q}}$ is a category of open modalities on a quantale $\mathcal{Q}$.
  Subexponential interpretation is a contravariant functor $\sigma : I \to \Box_{\mathcal{Q}}$ defined as follows:

  $\sigma(s_i) = \begin{cases}
  \Box_i : \mathcal{Q} \to \mathcal{Q} \text{, s.t.} \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a\},
  \\ \:\:\:\: \text{if $s_i \notin W \cap C \cap E$} \\
  \Box_i : \mathcal{Q} \to \mathcal{Q} \text{, s.t.} \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a, s \leq \mathds{1}\},
  \\ \:\:\:\: \text{if $s_i \in W$} \\
  \Box_i : \mathcal{Q} \to \mathcal{Q} \text{, s.t.} \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a, s \in \mathcal{Z}(\mathcal{Q}) \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  \Box_i : \mathcal{Q} \to \mathcal{Q} \text{, s.t.} \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a, \forall b, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  \text{otherwise, if $s_i$ belongs to some intersection of subsets, then we combine the relevant conditions } \\
  \end{cases}$
\end{defin}

@@

@@

\begin{proof}
We consider the promotion case, the rest modal cases are immediatly shown.

Let $!_{s_1} A_1, \dots, !_{s_n} A_n \rightarrow A$ and $\forall i, s \preceq s_i$.

Then $\forall a \in Q, \sigma(s_i)(a) \leq \sigma(s)(a)$.

By IH, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq [\![A]\!]$.

Thus, $\sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!]) \leq \sigma(s)([\![A]\!])$.

By Lemma 5, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq \sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!])$.

@@

\section{Quantale completeness}

@@

\begin{lemma}
Let $!_s \in I$ and $A \mathcal{F}_{\Sigma}$, then $\Box_s (\bigvee \{ A \}) = \bigvee \{ !_s B \: | \: !_s B \rightarrow A \}$
  is a quantic conucleus.
\end{lemma}

@@

See Yetter.
\end{proof}

\begin{lemma}
  Let $A \in \mathcal{F}_{\Sigma}$, then $\Box_s \bigvee \{ A \} = \bigvee \{ !_s A \}$, for each $s \in \mathcal{I}$.
\end{lemma}

\begin{proof}
  Let $A \in Fm$ and $s \in \mathcal{I}$.

@@

>>>>

\section{Introduction}

Categorial grammars were initially introduced by Ajdukiewicz and Bar-Hillel \cite{Ajd} \cite{Bar}.
In the 1950s, Lambek proposed the way of proof-theoretical consideration of such grammars \cite{Lambek}. In this approach, language is logic and parsing is derivation via inference rules. Let us consider a quite simple example. One needs to parse this quote from the poem by Oscar Wilde called "Impression du Matin" \cite{Wilde}:

\begin{center}
The Thames nocture of blue and gold \\
Changed to Harmony in grey
\end{center}

The first is to assign the correpsonding syntactic caterogies to the words as follows:

\vspace{\baselineskip}

$ $ \:\:\:\: The \:\:\:\: Thames nocturne \:\:\:\:\:\: of \:\:\:\:\:\:\:\:\:\: blue \:\:\:\:\:\:\: and \:\:\:\:\:\:\:\: gold

$(np / n) / np$ \:\:\:\:\: $np$ \:\:\:\:\:\:\:\:\: $n$ \:\:\:\:\:\: $np \backslash (np / ad)$ \: $ad$ \:\: $ad / (ad \backslash ad)$ \: $ad$

\vspace{\baselineskip}

$ $ \: Changed \:\:\:\: to \:\:\:\: Harmony \:\:\:\:\:\:\:\:\:\: in \:\:\:\:\:\:\:\:\:\:\: grey

$ $ \: $np \backslash (s / p)$ \:\: $p / np$ \:\:\:\:\:\: $np$ \:\:\:\:\:\:\:\: $np \backslash (np / ad)$ \:\: $ad$ \:\:\:\:\:\: $\rightarrow s$

\vspace{\baselineskip}

Here the basic reduction rules are:

\begin{itemize}
  \item $A, A \backslash B \to B$
  \item $B / A, A \to B$
\end{itemize}

Here we assign the special syntactic categories to the words of this sentence. $np$ denotes "noun phrase", $n$ -- noun, $ad$ -- adjective, $p$ -- phrase, $s$ -- sentence. This sequent denotes that this Oscar Wilde's quote is a well-formed sentence.
The verb "changed" has type "$np$ under ($s$ over $p$)". In other words, one needs to apply some noun phrase ("The Thames
nocturne of blue and gold") from the left, apply some phrase from the right ("Changed to Harmony in grey") and obtain
sentence after that. The other syntactic categories might be considered similarly.

The general case of such derivations in categorial grammars is axiomatised via Lambek calculus \cite{Lambek}, non-commutative
linear logic.

  \begin{prooftree}
    \AxiomC{$\Gamma, A, B, \Delta \rightarrow C$}
    \RightLabel{$\bullet \rightarrow$}
    \UnaryInfC{$\Gamma, A \bullet B, \Delta \rightarrow C$}
  \end{prooftree}
\end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}
  \begin{prooftree}
    \AxiomC{$A, \Pi \rightarrow B$}
    \RightLabel{$\rightarrow \backslash, \Pi \text{ is non-empty}$}
    \UnaryInfC{$\Pi \rightarrow A \backslash B$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Pi, A \rightarrow B$}
    \RightLabel{$\rightarrow /, \Pi \text{ is non-empty}$}
    \UnaryInfC{$\Pi \rightarrow B / A$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow A$}
    \AxiomC{$\Delta \rightarrow B$}
    \RightLabel{$\rightarrow \bullet$}
    \BinaryInfC{$\Gamma, \Delta \rightarrow A \bullet B$}
  \end{prooftree}
\end{flushright}
\end{minipage}

\begin{prooftree}
  \AxiomC{$\Gamma \rightarrow A$}
  \AxiomC{$\Pi, A, \Delta \rightarrow B$}
  \RightLabel{${\bf cut}$}
  \BinaryInfC{$\Gamma, \Pi, \Delta \rightarrow B$}
\end{prooftree}

Note that, non-commutativity is no coincidence. We cannot assume that a sentence remains well-formed regardless
of the word order.

As you could see, non-commutativity yields two implications: the left one (or, left division) and the right one (or, right division). Moreover, the right introduction rules for these divisions have the special restriction that claims non-emptiness
of the sequence $\Pi$ in the antecedent. This restriction denotes that we have no empty word that makes no sense
linguistically. On the other hand, we can avoid this restriction. This calculus is called $L^{*}$. Note that, the empty word
might be introduced explicilty via the constant ${\bf 1}$. The logic $L_{\bf 1}$ is a conservative extension of $L^{*}$ with
the following inference rules:

\begin{minipage}{0.5\textwidth}
\begin{flushleft}

  \end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}

\end{flushright}
\end{minipage}

Also, we have a multiplicative conjunction or non-commutative tensor. This connective plays the role of concatenation.

One may extend Lambek calculus with so-called additive conjunction and disjunction. The inference rules for these connections
are quite similar to inference rules for conjunction and disjunction in intuitionistic sequent calculus. In $L$-models,
additive connections are intersection and union of languages:

\begin{minipage}{0.5\textwidth}
\begin{flushleft}
    \begin{prooftree}
      \AxiomC{$\Gamma, A_i, \Delta \rightarrow B$}
      \RightLabel{$\land \rightarrow, i = 1,2$}
      \UnaryInfC{$\Gamma, A_1 \land A_2, \Delta \rightarrow B$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma, A, \Delta \rightarrow C$}
      \AxiomC{$\Gamma, B, \Delta \rightarrow C$}
      \RightLabel{$\vee \rightarrow$}
      \BinaryInfC{$\Gamma, A \vee B, \Delta \rightarrow C$}
    \end{prooftree}
  \end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}
  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow A$}
    \AxiomC{$\Gamma \rightarrow B$}
    \RightLabel{$\rightarrow \land$}
    \BinaryInfC{$\Gamma \rightarrow A \land B$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow A_i$}
    \RightLabel{$\rightarrow \vee, i = 1,2$}
    \UnaryInfC{$\Gamma \rightarrow A_1 \vee A_2$}
  \end{prooftree}
\end{flushright}
\end{minipage}

@@

Let us consider the following phrase as an example.

\begin{center}
  The young lady whom Childe Harold met before his pilgrimage
\end{center}

There is "the young lady" in the middle of this phrase. We are incapable of processing such cases in the core Lambek
calculus. That is, there is no ability to extract from the middle. "Met" is a phrasal verb. Whom did Childe Halord meet? He
met the young lady.

  \vspace{\baselineskip}

\begin{center}
The young lady whom$_i$ Childe Harold met $e_i$ before his pilgrimage
\end{center}

  \vspace{\baselineskip}

For this purpose, we introduce the exponential modality with exchange rule. The left one rule is a
dereliction, which is similar to the left $\Box$ introduction in modal logic $T$.

  \vspace{\baselineskip}

\begin{minipage}{0.5\textwidth}
\begin{flushleft}
  \begin{prooftree}
    \AxiomC{$\Gamma, A, \Delta \rightarrow B$}
    \RightLabel{$(! \to)$}
    \UnaryInfC{$\Gamma, ! A, \Delta \rightarrow B$}
  \end{prooftree}
\end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}
  \begin{prooftree}
    \AxiomC{$\Gamma, ! A, \Delta, \Theta \rightarrow C$}
    \RightLabel{${\bf ex}$}
    \UnaryInfC{$\Gamma, \Delta, ! A, \Theta \rightarrow C$}
  \end{prooftree}
\end{flushright}
\end{minipage}

  \vspace{\baselineskip}

The another one phrase is related to so-called parasitic extraction:

\begin{center}
The letter that Young Werther sent to Charlotte without reading
\end{center}

In addition to medial extraction, we used "the letter" twice in this phrase. Thus, one need to multiply our linguistical resources in a restricted way.

\begin{center}
The letter that$_i$ Young Werther sent $e_i$ to Charlotte without reading $e_i$
\end{center}

Subexponential modality with non-local contraction allows one to do such operations:

  \vspace{\baselineskip}

\begin{minipage}{0.5\textwidth}
\begin{flushleft}
  \begin{prooftree}
    \AxiomC{$\Gamma, ! A, \Delta, ! A, \Theta \rightarrow B$}
    \RightLabel{${\bf contr}$}
    \UnaryInfC{$\Gamma, \Delta, ! A, \Theta \rightarrow B$}
  \end{prooftree}
\end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}
  \begin{prooftree}
    \AxiomC{$\Gamma, ! A, \Delta, ! A, \Theta \rightarrow B$}
    \RightLabel{${\bf contr}$}
    \UnaryInfC{$\Gamma, ! A, \Delta, \Theta \rightarrow B$}
  \end{prooftree}
\end{flushright}
\end{minipage}

  \vspace{\baselineskip}

Note that, the usual form of contraction yields the cut inadmissibility in contrast to non-local contraction, which generalises the contraction.

  \vspace{\baselineskip}

It is useful to have many modalities and distinguish them in accordance with their abilities. In other words, we are going to
consider the polymodal case. Let us introduce a subexponential signature, which is a preorder with upwardly closed subsets,
where $W$ denotes weakening, etc. Note that, a subexponential signature might have any cardinality you prefer: finite or
infinite.

@@

  $\Sigma = \langle \mathcal{I}, \preceq, \mathcal{W}, \mathcal{C}, \mathcal{E} \rangle$,
where $\langle \mathcal{I}, \preceq \rangle$ is a preorder.
$\mathcal{W}, \mathcal{C}, \mathcal{E}$ are upwardly closed subsets of $I$ and $\mathcal{W} \cap \mathcal{C} \subseteq \mathcal{E}$.
\end{defin}

The last condition claims that if there are weakening and contraction, then one may also exchange as follows.

\begin{prooftree}
  \AxiomC{$\Gamma, ! A, \Delta, \Theta \rightarrow B$}
  \RightLabel{${\bf weak}$}
  \UnaryInfC{$\Gamma, ! A, \Delta, ! A, \Theta \rightarrow B$}
  \RightLabel{${\bf ncontr}$}
  \UnaryInfC{$\Gamma, \Delta, ! A, \Theta \rightarrow B$}
\end{prooftree}

@@

    \vspace{\baselineskip}

\begin{minipage}{0.5\textwidth}
  \begin{flushleft}
    \begin{prooftree}
    \AxiomC{$\Gamma, A, \Delta \rightarrow C$}
    \RightLabel{$! \rightarrow$}
    \UnaryInfC{$\Gamma, !^{s} A, \Delta \rightarrow C$}
    \end{prooftree}

    \begin{prooftree}
    \AxiomC{$\Gamma, !^{s} A, \Delta, !^{s} A, \Theta \rightarrow B$}
    \RightLabel{${\bf ncontr}_1, s \in C$}
    \UnaryInfC{$\Gamma, !^{s} A, \Delta, \Theta \rightarrow B$}
    \end{prooftree}

@@

  \end{flushleft}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \begin{flushright}
    \begin{prooftree}
    \AxiomC{$!^{s_1} A_1, \dots, !^{s_n} A_n \rightarrow A$}
    \RightLabel{$\rightarrow !, \forall j, s_j \succeq s$}
    \UnaryInfC{$!^{s_1} A_1, \dots, !^{s_n} A_n \rightarrow !^{s} A$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma, !^{s} A, \Delta, \Theta \rightarrow B$}
      \RightLabel{${\bf ex}_2, s \in E$}
      \UnaryInfC{$\Gamma, \Delta, !^{s} A, \Theta \rightarrow A$}
    \end{prooftree}
  \end{flushright}
\end{minipage}

@@

\end{defin}

After that, we introduce inference polymodal inference rules for non-commutative subexponentials. One can apply substructural
rules only if there is the relevant index on the current modality. Also, the right introduction rule is a sort of the
generalised right $\Box$-introduction a la $K4$. Modality $!_s$ might be introduced on the right only if its index is less than any other subexponential index from the antecedent. That is, if we have already used stronger modality, then we may
apply the weaker one.

  \vspace{\baselineskip}

Let us consider the current proof-theoretical and algorithmic results on Lambek calculus with additives and subexponentials.
The first of all, the cut rule is admissible. Generally, this calculus is undecidable, but the fragment without non-local
contraction is PSPACE-complete. These results were obtained by Kanovich, Kuznetsov, Nigam, and Scedrov.

@@

\section{Quantale background}

Now we introduce a reader to quantales quite briefly. For more details, one may take a look at these books \cite{Quantale18}
\cite{Rosenthal}. The initial idea of considering quantales in linear logic belongs to Yetter \cite{Yetter}.

@@

  A quantale is a triple $\mathcal{Q} = \langle A, \bigvee, \cdot \rangle$, where $\langle A, \bigvee \rangle$
is a complete join semilattice and $\langle A, \cdot \rangle$ is a semigroup such that for each indexing set $I$:

@@

@@

Residuality property for those divisions holds straightforwardly:

\begin{center}
  $b \leq a \backslash c \Leftrightarrow a \cdot b \leq c \Leftrightarrow a \leq b / c$
\end{center}

@@

  If $\mathcal{Q}_1$, $\mathcal{Q}_2$ are unital quantales, then a unital homomorphism if a quantale homomorphism such that $f(\varepsilon) = \varepsilon$.
\end{defin}

@@

  Let $\mathcal{Q} = \langle A, \bigvee, \cdot \rangle$ be a quantale. $\mathcal{S} \subseteq \mathcal{Q}$ is said to be a subquantale, if $\mathcal{S}$ is closed under multiplication and joins.
\end{defin}

@@

Then, if $f : \mathcal{Q}_1 \to \mathcal{Q}_2$ is a quantale homomorphism, then $f(\mathcal{S}) \subseteq \mathcal{Q}_2$ is a subquantale of $\mathcal{Q}_2$.
\end{prop}

@@

  Follows from the same statements for semigroup and semilattice homomorphisms.
\end{proof}

@@

We introduce the notion of quantic conucleus, an interior operation on quantale, which is a lax homomorphism with respect to semigroup multiplication. For unital quantales, this conucleus preserves a neutral.

@@

  An open modality (or quantic conucleus) on quantale $\mathcal{Q}$ is a map $I : \mathcal{Q} \to \mathcal{Q}$, such that

\begin{enumerate}
  \item $I x \leq x$;
  \item $I x = I^2 x$;
  \item $x \leq y \Rightarrow I x \leq I y$;
  \item $I x \cdot I y = I (I x \cdot I y)$.
\end{enumerate}

For unital quantale, we require that $I e = e$.
\end{defin}

Here and below, $I^2 x$ denotes $I (I x)$.

Note that, one may replace the last condition on the one $I x \cdot I y \leq I (x \cdot y)$. Thus, a quantic conuleus is the special case of lax monoidal monad from a category-theoretic point of view.

@@

  We define a partial order on quantic conuclei on $\mathcal{Q}$ as
  $I_1 \leq I_2 \Leftrightarrow \forall a \in \mathcal{Q}, I_1 a \leq I_2 a$.
\end{defin}

\begin{lemma}
  $I_1 a_1 \cdot I_2 a_2 \leq I (I_1 a_1 \cdot I_2 a_2)$, where $I_i \leq I, i = 1,2$.
\end{lemma}

@@

  $I_1 a_1 \cdot I_2 a_2 \leq I_1 (I_1 a_1) \cdot I_2 (I_2 a_2) \leq I (I_1 a_1) \cdot I (I_2 a_2) \leq I (I_1 a_1 \cdot I_2 a_2)$
\end{proof}

Let us define the special cases of quantic conuclei that it would be useful for our purpose:

\begin{defin}
$ $
\begin{enumerate}
  \item A quantic conucleus is called central, if for all
  $a, b \in \mathcal{Q}, \Box a \cdot b = b \cdot \Box a$.

  \item A quantic conucleus is called strong square-increasing,
  if for all $a, b \in \mathcal{Q}, \Box a \cdot b \leq \Box a \cdot b \cdot \Box a$ and
  $b \cdot \Box a \leq \Box a \cdot b \cdot \Box a$.

  \item A quantic conucleus is called unital, if $\forall a \in Q, \Box a \leq e$.
\end{enumerate}
\end{defin}

Note that centrality and strong square-increasing yield centrality.

@@

  Let $I$ be an open modality on some unital quantale $\mathcal{Q} = \langle A, \bigvee, \cdot, e \rangle$.
  Then, if $I$ is unital and weak idempotent, then $I$ is central.
\end{lemma}

@@

  $\begin{array}{lll}
  & b \cdot I a \leq & \\
  & \:\:\:\:\:\: \text{Right strong square-increase}& \\
  &I a \cdot b \cdot I a \leq & \\
  & \:\:\:\:\:\: \text{Unitality}& \\
  & I a \cdot b \cdot e \leq & \\
  & \:\:\:\:\:\: \text{Identity}& \\
  &I a \cdot b \leq & \\
  & \:\:\:\:\:\: \text{Left strong square-increase}& \\
  &I a \cdot b \cdot I a \leq & \\
  & \:\:\:\:\:\: \text{Unitality}& \\
  &e \cdot b \cdot I a \leq & \\
  & \:\:\:\:\:\: \text{Identity}& \\
  &b \cdot I a&
  \end{array}$

Hence, $b \cdot I a = I a \cdot b$, so for all $a \in \mathcal{Q}, I a \in \mathcal{Z}(\mathcal{Q})$.

@@

  Let $\mathcal{Q}$ be a quantale and $S \subseteq \mathcal{Q}$ a subquantale,
  then $I : \mathcal{Q} \to \mathcal{Q}$, such that $\Box a = \bigvee \{ s \in S \: | \: s \leq a \}$, is an open modality.
\end{prop}

@@

We define a subexponential interpretation in a quite sophisticated manner. Suppose we have a contravariant map from signature
to the set of subquantales. Here, contravariance denotes that the stronger subexponential index (in sense of preorder) maps
to the weaker subquantale. The second one function maps a subquantale to its quantic conucleus according to the previous
proposition. We match the index pursuant to its subset. In other words, if $s \in W$, then the result of subexponential
interpretation is an unital quantic conucleus, etc. More formally:

@@

Let $\mathcal{Q}$ be a quantale and $\Sigma = \langle \mathcal{I}, \preceq, \mathcal{W}, \mathcal{C}, \mathcal{E} \rangle$ be
a subexponential signature. Suppose, we have $S : \Sigma \to Sub(\mathcal{Q})$, a contravariant map from this subexponential
signature to the set of subquantales of $\mathcal{Q}$, and map $M : \mathcal{P}(\mathcal{Q}) \to \Box_{\mathcal{Q}}$, where
$\Box_{\mathcal{Q}}$ is the set of quantic conuclei on $\mathcal{Q}$. Thus, a subexponential interpretation is the map
$\sigma = M \circ S$ such that:

  $\sigma(s_i) = \begin{cases}
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a\},
  \\ \:\:\:\: \text{if $s_i \notin W \cap C \cap E$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a, s \leq \mathds{1}\},
  \\ \:\:\:\: \text{if $s_i \in W$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a, s \in \mathcal{Z}(\mathcal{Q}) \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  I_i : \mathcal{Q} \to \mathcal{Q} \text{ s.t. } \forall a \in Q, \Box_i(a) = \{ s \in S_i \: | \: s \leq a, \forall b, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s \},
  \\ \:\:\:\: \text{if $s_i \in E$} \\
  \text{otherwise, if $s_i$ belongs to some intersection of subsets, then we combine the relevant conditions } \\
  \end{cases}$
\end{defin}

@@

An interpretation and an entailment relation are defined standardly via valuation map and subexponential interpretation. There occur soundness and completeness theorems:

@@

We consider the promotion case, the rest modal cases are immediatly shown.

Let $!_{s_1} A_1, \dots, !_{s_n} A_n \rightarrow A$ and $\forall i, s \preceq s_i$. Then $\forall a \in Q, \sigma(s_i)(a) \leq \sigma(s)(a)$.

By IH, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq [\![A]\!]$. Thus, $\sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!]) \leq \sigma(s)([\![A]\!])$.

By Lemma, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq \sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!])$.

@@

Now we prove completeness a la MacNeille completetion. As a matter of fact, we generalised the technique provided by Brown and Gurr for the polymodal case.

@@

\begin{enumerate}
\item Let $!_s \in I$ and $A \mathcal{F}_{\Sigma}$, then $\Box_s (\bigvee \{ A \}) = \bigvee \{ !_s B \: | \: !_s B \rightarrow A \}$ is a quantic conucleus
\item Let $A \in \mathcal{F}_{\Sigma}$, then $\Box_s \bigvee \{ A \} = \bigvee \{ !_s A \}$, for each $s \in \mathcal{I}$
\item  Let $i, j \in I$ and $i \preceq j$, then for all $A \in \mathcal{F}_{\Sigma}$, $\Box_j (\bigvee \{ A \}) \subseteq \Box_i (\bigvee \{ A \})$.
\end{enumerate}

\end{lemma}

@@

\begin{enumerate}
\item Similarly to [Yetter].
\item Let $A \in Fm$ and $s \in \mathcal{I}$.

@@

\section{Categorical part}
